Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-09 VegOzone.tex
Refine intro to section 3.1

Data from before the 1990s are available solely from Svanvik and Karasjok. These data, however, do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}. Especially, ozone monitors did not undergo regular re-calibration. This probably leads to drifts in the observed data and may impose a false periodicity. In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjokka to a site close by Karasjok. Although the same equipment has been used, this might have introduced a baseline shift in the combined data series, denoted as Jergul/Karasjok, which would be particularly problematic if these were the sole data for comparison. Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are listed for completeness. There is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik (Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}).

@@

\section{Statistical analysis}
\label{sec:stats}
In this section, we assess the climate conditions. To this end, we compute multi-annual mean (referred to as climatology) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd. We evaluate the statistical significance of divergances from the norm in these variables for the growing seasons 2018/19 (referred to as anomalies) and perform a bias correction on the ozone climatology at Svanvik accounting for the increase in tropospheric background \chem{[O_3]}. For this, we cross-calibrate with long-term observations of \chem{[O_3]} at the sites in northern Finland and Sweden.

>>>>

Data from before the 1990s are available solely from Svanvik and Jergul. These data, however, do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}. Especially, ozone monitors did not undergo regular re-calibration. This probably leads to drifts in the observed data and may impose a false periodicity. In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjokka to a site close by Karasjok. Although the same equipment has been used, this might have introduced a baseline shift in the combined data series, denoted as Jergul/Karasjok, which would be particularly problematic if these were the sole data for comparison. Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are listed for completeness. There is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik (Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}).

@@

\section{Statistical analysis}
\label{sec:stats}
In this section, we assess the climate conditions. To this end, we compute multi-annual mean (referred to as climatology in the following) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd. For the growing seasons 2018/19, we evaluate the statistical significance of divergances from the norm in these variables (referred to as anomalies). We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Pallas and Esrange.
